D3064R0: Bringing the Laws of Physics, Computer-System Constraints, and Compilers to Bear on the OOTA Problem

### 2024-01-28 ootavacuous.tex
January 27 application of Alan Stern feedback, #2
Signed-off-by: Paul E. McKenney <paulmck@kernel.org>

@@

Given all of these complications in C++ executions, and given that
semantic dependencies must be defined in terms of executions, it makes
sense to lower our sights from full generality.
We therefore restrict ourselves to a specific context:
traditional compilers.
Not only that, we will take the notion of semantic dependence (and thus
OOTA) to be per-compiler rather than absolute.

For the special case of non-volatile atomic operations, we will further
restrict ourselves to compilers that do single-threaded analysis, with
the possible addition of specific constraints deduced from multi-threaded
analysis.
For volatile atomic operations, compilers are prohibited from inventing
loads, and we will restrict ourselves to compilers that extend that
prohibition to non-volatile atomic loads.

>>>>

This definition makes OOTA more complicated than a simple cycle but
we will continue to refer to it as an OOTA cycle out of habit.
Note that this new definition includes and generalizes our previous
\co{acyclic(sdep ; rfe)} definition.

@@

Given all of these complications in C++ executions, and given that
semantic dependencies must be defined in terms of executions, it makes
sense to lower our sights from full generality.
We therefore restrict ourselves to a specific context:
traditional compilers.
It will then turn out to be necessary to take the notion of semantic
dependence, and thus OOTA, to be compiler-relative rather than absolute.

We will show that programs built using strict C++ compilers that treat all
atomic operations as volatile can never exhibit OOTA, and the same is true
for compilers obeying the weaker constraint that atomic accesses may not
invented, provided the compiler does not perform multi-thread analysis.
